Automated Verification and Synthesis of Cyber-Physical Systems: Advances and Challenges

### 2018-03-25 IET CPS/SurveyECPS.tex
IET CPSTA

\documentclass[format=acmsmall, review=false, screen=true]{acmart}

@@

\newtheorem{prf}{\textbf{Definition}}

@@

% Metadata Information
%\acmJournal{TCPS}
%\acmVolume{9}
%\acmNumber{9}
%\acmArticle{99}
%\acmYear{2017}
%\acmMonth{12}
%\copyrightyear{2017}
%\acmArticleSeq{9}

% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{0000001.0000001}

% Paper history
\received{August 2017}
%\received[revised]{March 2009}
%\received[accepted]{June 2009}

\lstset{language=C,basicstyle=\small}
\lstset{numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt}
\lstset{tabsize=2}
\lstset{firstnumber=1}
\lstset{frame=single}
\lstset{
  language={C},
  morekeywords={stdio,math,printf,assume,then,end-if,end,input,output}
}

\usepackage{color}
\usepackage{listings}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
%\renewcommand{\baselinestretch}{0.93}
\renewcommand{\baselinestretch}{0.97}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Automated Verification and Synthesis of Cyber-Physical Systems: Advances and Challenges}

\author{Lucas C. Cordeiro}
\orcid{0000-0002-6235-4272}
\affiliation{%
  \institution{Department of Computer Science, University of Oxford}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Oxford}
%  \state{VA}
%  \postcode{23185}
  \country{UK}}

\author{Eddie B. de Lima Filho}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{TPV Technology}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

\author{Iury V. de Bessa}
\orcid{0000-0002-2758-4638}
\affiliation{%
  \institution{Department of Electricity, Federal University of Amazonas}
%  \streetaddress{Wolfson Building, Parks Rd, Oxford OX1 3QD}
  \city{Manaus}
%  \state{VA}
%  \postcode{23185}
  \country{Brazil}}

%% author names and affiliations
%% use a multiple column layout for up to three different
%% affiliations
%\author{Eddie B. de Lima Filho \\
%%Department of Computer Science\\
%University of Oxford, UK\\
%lucas.cordeiro@cs.ox.ac.uk
%\and
%Eddie B. de Lima Filho \\
%Samsung Electronics, Brazil \\
%eddie.l@samsung.com
%\and
%Iury V. de Bessa \\
%Federal University of Amazonas, Brazil \\
%iury.bessa@ufam.edu.br}
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678--2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% make the title area

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors. Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPSs). As a consequence, reliability of ECPSs becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for ECPSs generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs. The present survey discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs. Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

%\IEEEpeerreviewmaketitle
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010405.10010432.10010439.10010440</concept_id>
<concept_desc>Applied computing~Computer-aided design</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010553</concept_id>
<concept_desc>Computer systems organization~Embedded and cyber-physical systems</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010178.10010213.10010214</concept_id>
<concept_desc>Computing methodologies~Computational control theory</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10002944.10011122.10002945</concept_id>
<concept_desc>General and reference~Surveys and overviews</concept_desc>
<concept_significance>100</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Computer-aided design}
\ccsdesc[500]{Computer systems organization~Embedded and cyber-physical systems}
\ccsdesc[300]{Computing methodologies~Computational control theory}
\ccsdesc[100]{General and reference~Surveys and overviews}

\keywords{Embedded Systems, Cyber-Physical Systems, Formal Methods, Model Checking, Machine Learning}

@@

Figure~\ref{intelligent-product} shows some examples of ECPSs, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}. Indeed, many current ECPSs, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences. On the one hand, portable ECPSs are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics. On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health. 
%
\begin{figure}[!t]
	\centering
	\includegraphics[width=0.6\textwidth]{figure1.jpg}
	\caption{An ECPS is part of a well-specified larger one (intelligent product).}
	\label{intelligent-product}
\end{figure}

@@

Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may change based on data acquired during previous actions, due to model updating. As a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data, as recently reported in the literature~\cite{DBLP:conf/cav/HuangKWW17,DBLP:conf/cav/KatzBDJK17}, where formal verification techniques are applied to verify deep neural networks.

@@

The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{achievements} tackles those challenges that are (partially) open in current published research and later describes current achievements and future trends in verification and synthesis of ECPSs, and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here. Finally, we conclude and describe future work, in Section~\ref{conclusions}.

@@

Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as provide controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}.

An ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho... parece que estamos misturando tudo... sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo um pouco nebuloso?}
The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

%--------------------------------------
\subsection{A Model for Cyber-physical Systems}
\label{ssec:model}
%--------------------------------------

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled, and they should be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

@@

Finally, in addition to ECPS-dynamics modeling, it is often interesting to establish invariants on a ECPS's state space, which can be used for accelerating formal verification and synthesis processes regarding those systems. Such invariants might be synthesized, in order to point out state-space boundaries~\cite{BenSassi12} and determine ECPS stability~\cite{Paul14} or reachability~\cite{Li16_2,Li16}.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region. Stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification} or linear system stability theory for discrete simulations~\cite{Bessa17}.

@@

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating physical-system state, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

@@

Bounded Model Checking (BMC) based on SAT was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}. 
%
In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{figure2.jpg}
	\caption{Bounded Model Checking.}
	\label{bounded-model-checking}
\end{figure}

@@

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). The SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable state has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.4\textwidth}
\begin{lstlisting}[basicstyle=\footnotesize]
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.4\textwidth}
\begin{lstlisting}[basicstyle=\footnotesize]
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

@@

\begin{prf}
\begin{definition}\label{interpolants}
\textit{\textcolor{blue}{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties}~\cite{McMillan03,McMillan05}: }

\begin{itemize}
	\item $A \Rightarrow F$;
	\item $F \wedge B$ \textit{is unsatisfiable};
	\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
\end{itemize}
\end{definition}
\end{prf}

@@

\textcolor{blue}{The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.}
%
\textcolor{blue}{In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries}, where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$. Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).}
%
\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{interpolants}
\caption{Computing image by interpolation~\cite{McMillan03}.}
\label{figure:computing-image-by-interpolation}
\end{figure}

>>>>

\documentclass{cta-author}

\newtheorem{theorem}{Theorem}{}
\newtheorem{corollary}{Corollary}{}
\newtheorem{remark}{Remark}{}
\newtheorem{definition}{Definition}{}

@@

%\newtheorem{prf}{\textbf{Definition}}

@@

\begin{document}

%\supertitle{Brief Paper}

\title{Automated Verification and Synthesis of Cyber-Physical Systems: Advances and Challenges}

\author{\au{Lucas C. Cordeiro$^{1\corr}$} \au{Eddie B. de Lima Filho$^2$} \au{Iury V. de Bessa$^3$}}

\address{\add{1}{Department of Computer Science, University of Oxford, Oxford, United Kingdom}
\add{2}{TPV Technology, Manaus, Brazil}
\add{3}{Department of Electricity, Federal University of Amazonas, Manaus, Brazil}
\email{lucasccordeiro@gmail.com}}

\begin{abstract}
\looseness=-1 Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors. Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPSs). As a consequence, reliability of ECPSs becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for ECPSs generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs. The present survey discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs. Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

@@

{\color{blue}
Figure~\ref{intelligent-product} shows some examples of ECPSs, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}. Indeed, many current ECPSs, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions and civil applications, since they can easily achieve places that cannot be accessed by humans without on-board pilots with different degrees of autonomy; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences. On the one hand, portable ECPSs are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics. On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health.
}
%
\begin{figure}[!t]
	\centering
	\includegraphics[width=\columnwidth]{figure1.jpg}
	\caption{An ECPS is part of a well-specified larger one (intelligent product).}
	\label{intelligent-product}
\end{figure}

@@

{\color{blue}
Regarding behavior correctness, one important observation is that it is becoming more difficult, since fixed mathematical models are being replaced by machine learning approaches (ML) \cite{mlembe}, where future operation may change based on data acquired during previous actions, due to model updating. Indeed, ML algorithms have a dual role on software verification development. With the dissemination of the use of these algorithms, several ECPS have embedded ML software, and as a consequence, system-correctness verification procedures should also adapt somehow and incorporate new checks, which may also rely on ML, when analyzing target models or fed with acquired data, as recently reported in the literature~\cite{DBLP:conf/cav/HuangKWW17,DBLP:conf/cav/KatzBDJK17}, where formal verification techniques are applied to verify deep neural networks. Otherwise, model checking mechanisms have also incorporated ML algorithms to increase the efficiency of verification~\cite{BUCCAFURRI199957,Bortolussi2015323,Braazdil201498} and synthesis~\cite{Jha2017693} procedures.
}

@@

{\color{blue}
The present article is an survey study whose main goals are described as follows:
\begin{itemize}
	\item briefly present part of the established research about ECPS modelling, verification and synthesis;
	\item provide an overview about the technical issues and advances in symbolic formal verification and synthesis methodologies that are currently employed for ECPS;
	\item present the open challenges on the ECPS verification and synthesis and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. 
\end{itemize}

The remaining of this survey is organized as follows. We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{achievements} tackles those challenges that are (partially) open in current published research and later describes current achievements and future trends in verification and synthesis of ECPSs, and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here. Finally, we conclude and describe future work, in Section~\ref{conclusions}.
}

@@

{\color{blue}

Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as provide controller synthesis~\cite{Nilsson16,Prabhakar17,EsmaeilZadehSoudjani17,Ames17,tabuada2009verification,Hasuo17,Zamani15} for ECPSs and HSs~\cite{Clarke11,Aerts17,Zhang14,Simko14,Sanwal13,Lee13,Li14}. The reliability of an ECPS is related to its ability to maintain the continuity of correct execution of specific function. In particular, the reliability of an ECPSs depends on the correctness of their embedded software and the compatibility with hardware. The correctness is asserted with respect to a given specification, that may be a safety specification or a reachability specification~\cite{Jhala:2009:SMC:1592434.1592438}. A safety specification is related to  a set of undesirable states that must be avoided by the system, and a reachability is related to a desired state, that must be achieved. In both cases, it is needed a model to compute the reachability of ECPS an than decide about the correctness and reliability.

A general model for a ECPS is defined as follows.

\begin{definition}
	An ECPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.
\end{definition}

%\textcolor{red}{confesso que, para mim, o paragrafo abaixo ficou estranho... parece que estamos misturando tudo... sera que na tengtativa de implementar os comentarios e ter o artigo aceito nao estamos deixando o artigo um pouco nebuloso?}

The synthesis of reliable ECPS is related to the synthesis of controllers that are able to handle the complex dynamic of these systems. Indeed, the so-called ECPS can be considered HS, since they present a mixture of continuous and discrete dynamics. Thus, the control synthesis for ECPS is based in the hybrid control system theory, that is still in development, since the control theory for continuous and discrete systems were developed without connexions~\cite{lunze_lamnabhi-lagarrigue_2009}. The formal synthesis of ECPS includes but is not limited to the control synthesis problem, and must consider the formal verification of correctness and reliability.

Summarizing, the overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

}

%The classification of approaches in terms of their deterministic vs stochastic nature should be definitely revised: a model is non-deterministic when it accepts different behaviors without providing a measure of probability (e.g. this is the case of Timed Automata, which are instead mentioned as a deterministic model); a model is stochastic when the space of behaviors is provided with a measure of probability. By the way, a large literature now develops on mixed models that combine non-deterministic and stochastic behavior (e.g. Markov Decision Processes), but this class of models is not even mentioned.

%--------------------------------------
\subsection{A Model for Cyber-physical Systems}
\label{ssec:model}
%--------------------------------------
{
\color{blue}

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness which is described by probability distributions, i.e., some of these variable are random variables whose behavior are described by probability density functions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled, and they should be better represented by nondeterministic approaches~\cite{leeCPS2}. Non-deterministic models is able to model different behaviors without providing a measure of probability, i.e., their variables are assumed to be different values but they are not random variables. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.
}

@@

{\color{blue}
In addition to ECPS-dynamics modeling, it is often interesting to establish invariants on a ECPS's state space, which can be used for accelerating formal verification and synthesis processes regarding those systems. Such invariants might be synthesized, in order to point out state-space boundaries~\cite{BenSassi12} and determine ECPS stability~\cite{Paul14} or reachability~\cite{Li16_2,Li16}. Finally, it is worth to mention that there are still some recent modelling approaches that are mixtures of stochastic and non-deterministic models for ECPS, e.g., stochastic nondeterministic automata~\cite{Fisher2014}, and piecewise-deterministic Markov chains~\cite{Julius09}.
}

@@

{\color{blue}

An ECPS verification problem is defined as follows:

\begin{definition}
	Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}.
\end{definition}

 Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region. Stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification} or linear system stability theory for discrete simulations~\cite{Bessa17}. 
}

@@

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating physical-system state, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{ARAUJO2017,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

@@

Bounded Model Checking (BMC) based on SAT was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}. 
%
In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\columnwidth]{figure2.jpg}
	\caption{Bounded Model Checking.}
	\label{bounded-model-checking}
\end{figure}

@@

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). The SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable state has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
	\centering
	\begin{minipage}{\columnwidth}
		\begin{lstlisting}[basicstyle=\footnotesize]
		#include <assert.h>
		int x[N], a;
		...
		int main(void) {
		a=N;
		for(int i=0; i<N; i++)
		if (x[i]>1)
		a--;
		assert(a<=N);
		return 0;
		}
		\end{lstlisting}
	\end{minipage}
	\begin{center}
		(a)
	\end{center}
	\centering
	\begin{minipage}{0.4\textwidth}
		\begin{lstlisting}[basicstyle=\footnotesize]
		a1 = N
		a2 = (x[0] > 1) ? a1 - 1 : a1
		a3 = (x[1] > 1) ? a2 - 1 : a2
		a4 = (x[2] > 1) ? a3 - 1 : a3
		...
		an+1 = (x[N-1] > 1) ? an - 1 : an 
		\end{lstlisting}
	\end{minipage}
	\begin{center}
		(b)
	\end{center}
	\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
	\label{figure:verification-condition}
\end{figure}

@@

%\begin{prf}
	\begin{definition}\label{interpolants}
		\textit{\textcolor{blue}{Given a pair of formulae $\left(A,B\right)$, such that $A \wedge B$ is inconsistent, and a proof by resolution for $\left(A,B\right)$, an \textit{interpolant} for $\left(A,B\right)$ is a formula $F$ with the following properties}~\cite{McMillan03,McMillan05}: }

		\begin{itemize}
			\item $A \Rightarrow F$;
			\item $F \wedge B$ \textit{is unsatisfiable};
			\item $F$ \textit{expressed only over the common variables (nonlogical symbols) of $A$ and $B$}.
		\end{itemize}
	\end{definition}
%\end{prf}

@@

\textcolor{blue}{The use of interpolants allows us to define a complete method for finite-state reachability analysis based on SAT and SMT solvers.}
%
\textcolor{blue}{In order to show how BMC and interpolation can be combined, we refer to Section~\ref{Preliminaries}, where we define Eq.~(\ref{bounded-model-checking-biere}) and the terms $I$, $R$, and $\phi$. Now, suppose that $Q=I$ and Eq.~(\ref{bounded-model-checking-biere}) is partitioned, so that the set of initial states $I$ and the first instance of the transition relation $R$ are in set $A$, while the remaining instances of $R$ and the property $\phi$ are in set $B$, as shown in Figure~\ref{figure:computing-image-by-interpolation} (note that $k$ is unknown).}
%
\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{interpolants}
	\caption{Computing image by interpolation~\cite{McMillan03}.}
	\label{figure:computing-image-by-interpolation}
\end{figure}
